Malacology: A Programmable Storage System

### 2016-10-15 paper/paper.tex
tweak
tweaks

%\usepackage{fullpage}
%\usepackage{times}
\usepackage{multirow}
\usepackage{arydshln}
%\usepackage{amssymb,amsmath}
%\usepackage{ifxetex,ifluatex}
\usepackage{amsmath}
%\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{listings}
%\usepackage{amssymb}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
%\newcommand{\xmark}{\ding{55}}%
%
%
%
%% use upquote if available, for straight quotes in verbatim environments
%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
%
%\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
%% Scale images if necessary, so that they will not overflow the page
%% margins by default, and it is still possible to overwrite the defaults
%% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,keepaspectratio}

@@

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
their code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to balance load, efficiently propagate cluster metadata, recover
from failure, and to make trade-offs for latency/throughput using leases.
\end{abstract}

@@

A number of recent developments have disrupted traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. 
Second, the demand for
managing structured data and flexible consistency semantics at scale pressure
big data processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}. 
Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

These three trends put evolutionary pressure on storage systems and raise the
question of whether there are principles that storage systems designers can follow
to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.
% that are important for
%users to continue to entrust their data to the storage system.

In this paper we investigate an approach that focuses on generalizing and
composing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. 
%We will refer to this programmability as
%\emph{programmable storage}, which differs from \emph{active storage} (the
%injection and execution of arbitrary code in a storage system) 
%and
%\emph{software-defined storage} (the control of thin-provisioning of storage).

{\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
re-use and extension of existing storage abstractions and services provided by the underlying software stack,
to enable the creation of new storage abstractions via composition.
A programmable storage system is realized by exposing existing functionality (such as 
metadata services and synchronization capabilities) as interfaces that can be ``glued together''
using a high-level language.  Programmable storage differs from \emph{active storage},
or the injection and execution of arbitrary code in a storage system.

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the Malacology in Ceph, a popular open source software storage stack.
We choose Ceph to demonstrate the concept of programmable storage because it
offers a broad spectrum of existing services, including distributed locking and
caching services provided by metadata servers, durability and object interfaces
provided by the backend object store, and propagation of consistent
cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

@@

{\it \textbf{We implement two distributed services}} using programmable storage
to demonstrate its feasability:

@@

Building storage stacks from the ground-up for a specific purpose results in
the best performance. For example, HDFS was designed specifically to serve
Hadoop's jobs, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
the best performance~\cite{CITEME}. Alternatively, general-purpose storage
stacks are built with the flexiblity to serve many applications by providing a
variety of interfaces and tunable parameters. Unfortunately, managing competing
forces in these systems is difficult and users
want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.

>>>>

%\usepackage{fullpage}
%\usepackage{times}
\usepackage{multirow}
\usepackage{arydshln}
%\usepackage{amssymb,amsmath}
%\usepackage{ifxetex,ifluatex}
\usepackage{amsmath}
%\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{listings}
%\usepackage{amssymb}
\usepackage{pifont}
\usepackage{color}
\newcommand{\cmark}{\ding{51}}%

\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
%\newcommand{\xmark}{\ding{55}}%
%
%
%
%% use upquote if available, for straight quotes in verbatim environments
%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
%
%\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
%% Scale images if necessary, so that they will not overflow the page
%% margins by default, and it is still possible to overwrite the defaults
%% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,keepaspectratio}

@@

\begin{abstract}
Storage systems are caught between rapidly changing data
processing applications and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to support rapid evolution both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
their code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to balance load, efficiently propagate cluster metadata, recover
from failure, and to make trade-offs for latency/throughput using leases.
\end{abstract}

@@

A number of recent developments have disrupted traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. 
Second, the demand for
managing structured data and flexible consistency semantics at scale pressure
big data processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}. 
\paa{the previous sentence doesn't scan well. do we just want to say that 
emerging ``big data'' applications demand interface evolution to support flexible
consistency as well as flexible structured data representation?}
Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

%These three trends put evolutionary pressure on storage systems and raise the
%question of whether 
The evolutionary pressure placed on storage systems by these trends raises the
question of whether
there are principles that storage systems designers can follow
to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.
% that are important for
%users to continue to entrust their data to the storage system.

In this paper we investigate an approach that focuses on 
%generalizing 
%and
%composing 
identifying and exposing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. 
%We will refer to this programmability as
%\emph{programmable storage}, which differs from \emph{active storage} (the
%injection and execution of arbitrary code in a storage system) 
%and
%\emph{software-defined storage} (the control of thin-provisioning of storage).

{\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
re-use and extension of existing storage abstractions provided by the underlying software stack,
to enable the creation of new services via composition.
A programmable storage system can be realized by exposing existing functionality (such as 
metadata services and synchronization and monitoring capabilities) as interfaces that can be ``glued together''
in a variety of ways
using a high-level language.  Programmable storage differs from \emph{active storage},
or the injection and execution of arbitrary code in a storage system.

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system 
that facilitates the construction of
new
services by re-purposing existing subsystem abstractions of the storage stack.
We build Malacology in Ceph, a popular open source software storage stack.
We choose Ceph to demonstrate the concept of programmable storage because it
offers a broad spectrum of existing services, including distributed locking and
caching services provided by metadata servers, durability and object interfaces
provided by the backend object store, and propagation of consistent
cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

@@

{\it \textbf{We implement two distributed services}} using 
%programmable storage
%to demonstrate its feasability:
Malacology to demonstrate the feasibility of the programmable storage approach:

@@

Building storage stacks from the ground-up for a specific purpose results in
the best performance. For example, HDFS was designed specifically to serve
Hadoop's jobs~\paa{lol not sure if I buy that hadoop achieves ``best performance''}, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
the best performance~\cite{CITEME}. Alternatively, general-purpose storage
stacks are built with the flexiblity to serve many applications by providing a
variety of interfaces and tunable parameters. Unfortunately, managing competing
forces in these systems is difficult and users
want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.
